\textbf{The Mint Scale: A Fresh Validation of the Multimodal Interoception Questionnaire and Comparison to the MAIA, BPQ and IAS}

### 2025-08-20 paper/manuscript.qmd
update

@@

Next, we tested the predictive power of the Mint scale relative to other interoception questionnaires.
We assessed the importance of each interoceptive dimension (from all the scales) as a unique predictor by fitting regression models (linear for continuous measures - e.g., depression score - and logistic for binary variables - e.g., presence *vs.* absence of mood disorders) to predict different outcomes, and compare the best between the 4 interoception scales based on the total explained variance (R2).
We then evaluated the predictive performance of each scale as a whole by comparing regression models with all the dimensions entered as predictors (note that the IAS and BPQ only have one total score variable). We assessed the models based on their R2, as well as on the Bayesian Information Criterion (BIC), which penalizes for predictor number, thus offering a balance between model parsimony and predictive power. For the logistic models, we quantified the discriminative power by computing the Area Under the Curve (AUC) of the Receiver Operating Characteristic (ROC) curve, which assesses the model's discriminative power (the combination of sensitivity and specificity).

>>>>

For both studies, data analysis was carried out using R [4.4.3, @Rcoreteam], leveraging the *easystats* ecosystem [@ludecke2019insight; @ludecke2020extracting] and the *EGAnet* package [@christensen2021estimating]. The reproducible code and additional details are available at https://github.com/RealityBending/InteroceptionScale.

@@

Next, we tested the predictive power of the Mint scale relative to other interoception questionnaires.
We assessed the importance of each interoceptive dimension (from all the scales) as a unique predictor by fitting regression models (linear for continuous measures - e.g., depression score - and logistic for binary variables - e.g., presence *vs.* absence of mood disorders) to predict different outcomes, and compare the best between the 4 interoception scales based on the total explained variance [R2 or pseudo-R2 for logistic models, @ludecke2021performance].
We then evaluated the predictive performance of each scale as a whole by comparing regression models with all the dimensions entered as predictors (note that the IAS and BPQ only have one total score variable). We assessed the models based on their R2, as well as on the Bayesian Information Criterion (BIC), which penalizes for predictor number, thus offering a balance between model parsimony and predictive power. For the logistic models, we quantified the discriminative power by computing the Area Under the Curve (AUC) of the Receiver Operating Characteristic (ROC) curve, which assesses the model's discriminative power (the combination of sensitivity and specificity).

### 2025-08-20 paper/manuscript.tex
update

\authornote{\par{\addORCIDlink{Dominique
Makowski}{0000-0001-5375-9967}}\par{\addORCIDlink{Ana
Neves}{0009-0006-0020-7599}}\par{\addORCIDlink{Giulia
Poreiro}{0000-0002-2343-5109}} 
\par{ }
\par{     \begin{tcolorbox}[enhanced jigsaw, breakable, colframe=quarto-callout-note-color-frame, leftrule=.75mm, arc=.35mm, colback=white, toprule=.15mm, bottomrule=.15mm, rightrule=.15mm, opacityback=0, left=2mm]

This preprint is a non-peer-reviewed work from the
\href{https://realitybending.github.io/}{\textbf{Reality Bending Lab}}.
\begin{center}
\includegraphics[width=0.2\linewidth,height=\textheight,keepaspectratio]{manuscript_files/mediabag/ReBeL_LogoOnly_hu114.png}
\end{center}

@@

These differing philosophies have practical consequences, leading to
poor convergence between scales that are assumed to measure the same
construct (\citeproc{ref-vig2022questionnaires}{Vig et al., 2022}) - a
classic example of the ``jingle fallacy'' where the shared label of
``interoception'' can mask underlying measures and differences,
hindering replication and theoretical integration. In acknowledgement of
this issue, efforts have been made to align subjective measures with
theoretical models that outline different facets of interoception (e.g.,
accuracy, attention, deficits). However, focusing on developing measures
that are theory-driven may however be premature when theoretical models
are in flux, and when a topographic mapping to self-report dimensions is
not guaranteed. For example, interoceptive distinctions that are
relevant at a neuroanatomical or cognitive level might not necessarily
map onto the structure of subjective experiences. More fundamentally,
existing interoception questionnaires typically fail to consider the
full range of interoceptive modalities and the context in which
interoceptive experiences occur. Most existing questionnaires focus on a
narrow range of bodily modalities, with a heavy bias towards
cardio-respiratory sensations at the expense of others like gastric,
urogenital, or thermoregulatory signals. Perhaps most critically,
existing measures ask participants to make general assessments across
most/all contexts, which does not capture the potential homeostatic
relevance of attending or detecting internal signals. Without a
contextual reference point, there is likely to be substantial
variability in how participants interpret and respond to items. For
example, a general item about ``feeling your heart beat'' is ambiguous
since its salience, frequency, experience and interpretation would
differ depending on the context, e.g., on whether it occurs during
physical exercise, a moment of anxiety, or quiet rest (e.g., someone who
exercises regularly, or is anxious, might more easily recall an instance
of occurrence of feeling their heart beating and respond based on that
contextual saliency).

@@

The structure of the items was analyzed using the recently-developed
Exploratory Graph Analysis (EGA, \citeproc{ref-golino2017exploratory}{H.
F. Golino \& Epskamp, 2017}) framework, which allows to jointly estimate
the number of dimensions (i.e., clusters of items), the structure, as
well as its stability using bootstrapping
(\citeproc{ref-golino2020investigating}{H. Golino et al., 2020}). At a
fundamental level, EGA conceptualizes variables as nodes in a network,
with connections (edges) reflecting associations between them. Evidence
has underlined its suitability as an alternative to traditional factor
analysis, addressing some of its limitations such as the assumption of a
``latent'' source of variability, issues in estimation of the optimal
factor numbers, and poor performance in complex population structures,
while remaining comparable and interpretable
(\citeproc{ref-christensen2021equivalency}{Christensen \& Golino, 2021};
\citeproc{ref-jimenez2023dimensionality}{Jiménez et al., 2023}). In
particular, nodes communities (i.e., clusters of items) can be in
practice interpreted as distinct ``dimensions'', similarly to
traditional latent factors - but without explicitly assuming their
existence (\citeproc{ref-christensen2021equivalency}{Christensen \&
Golino, 2021}).

@@

@@

Next, we tested the predictive power of the Mint scale relative to other
interoception questionnaires. We assessed the importance of each
interoceptive dimension (from all the scales) as a unique predictor by
fitting regression models (linear for continuous measures - e.g.,
depression score - and logistic for binary variables - e.g., presence
\emph{vs.} absence of mood disorders) to predict different outcomes, and
compare the best between the 4 interoception scales based on the total
explained variance (R2). We then evaluated the predictive performance of
each scale as a whole by comparing regression models with all the
dimensions entered as predictors (note that the IAS and BPQ only have
one total score variable). We assessed the models based on their R2, as
well as on the Bayesian Information Criterion (BIC), which penalizes for
predictor number, thus offering a balance between model parsimony and
predictive power. For the logistic models, we quantified the
discriminative power by computing the Area Under the Curve (AUC) of the
Receiver Operating Characteristic (ROC) curve, which assesses the
model's discriminative power (the combination of sensitivity and
specificity).

@@

\bibitem[\citeproctext]{ref-christensen2021equivalency}
Christensen, A. P., \& Golino, H. (2021). On the equivalency of factor
and network loadings. \emph{Behavior Research Methods}, \emph{53}(4),
1563--1580.

@@

@@

>>>>

\authornote{\par{\addORCIDlink{Dominique
Makowski}{0000-0001-5375-9967}}\par{\addORCIDlink{Ana
Neves}{0009-0006-0020-7599}}\par{\addORCIDlink{Giulia
Poreiro}{0000-0002-2343-5109}} 
\par{ }
\par{     \begin{tcolorbox}[enhanced jigsaw, breakable, opacityback=0, rightrule=.15mm, colback=white, toprule=.15mm, arc=.35mm, bottomrule=.15mm, colframe=quarto-callout-note-color-frame, left=2mm, leftrule=.75mm]

This preprint is a non-peer-reviewed work from the
\href{https://realitybending.github.io/}{\textbf{Reality Bending Lab}}.
\includegraphics[width=0.2\linewidth,height=\textheight,keepaspectratio]{manuscript_files/mediabag/ReBeL_LogoOnly_hu114.png}

@@

These differing philosophies have practical consequences, leading to
poor convergence between scales that are assumed to measure the same
construct (\citeproc{ref-vig2022questionnaires}{Vig et al., 2022}) - a
classic example of the ``jingle fallacy'' where the shared label of
``interoception'' can mask underlying measurand differences, hindering
replication and theoretical integration. In acknowledgement of this
issue, efforts have been made to align subjective measures with
theoretical models that outline different facets of interoception (e.g.,
accuracy, attention, deficits). However, focusing on developing measures
that are theory-driven may however be premature when theoretical models
are in flux, and when a topographic mapping to self-report dimensions is
not guaranteed. For example, interoceptive distinctions that are
relevant at a neuroanatomical or cognitive level might not necessarily
map onto the structure of subjective experiences. More fundamentally,
existing interoception questionnaires typically fail to consider the
full range of interoceptive modalities and the context in which
interoceptive experiences occur. Most existing questionnaires focus on a
narrow range of bodily modalities, with a heavy bias towards
cardio-respiratory sensations at the expense of others like gastric,
urogenital, or thermoregulatory signals. Perhaps most critically,
existing measures ask participants to make general assessments across
most/all contexts, which does not capture the potential homeostatic
relevance of attending or detecting internal signals. Without a
contextual reference point, there is likely to be substantial
variability in how participants interpret and respond to items. For
example, a general item about ``feeling your heart beat'' is ambiguous
since its salience, frequency, experience and interpretation would
differ depending on the context, e.g., on whether it occurs during
physical exercise, a moment of anxiety, or quiet rest (e.g., someone who
exercises regularly, or is anxious, might more easily recall an instance
of occurrence of feeling their heart beating and respond based on that
contextual saliency).

@@

The structure of the items was analyzed using the recently-developed
Exploratory Graph Analysis (EGA, \citeproc{ref-golino2017exploratory}{H.
F. Golino \& Epskamp, 2017}) framework, which allows to jointly estimate
the number of dimensions (i.e., clusters of items), the structure, as
well as its stability using bootstrapping
(\citeproc{ref-golino2020investigating}{H. Golino et al., 2020}). At a
fundamental level, EGA conceptualizes variables as nodes in a network,
with connections (edges) reflecting associations between them. Evidence
has underlined its suitability as an alternative to traditional factor
analysis, addressing some of its limitations such as the assumption of a
``latent'' source of variability, issues in estimation of the optimal
factor numbers, and poor performance in complex population structures,
while remaining comparable and interpretable
(\citeproc{ref-christensen2021equivalency}{Christensen \& Golino,
2021b}; \citeproc{ref-jimenez2023dimensionality}{Jiménez et al., 2023}).
In particular, nodes communities (i.e., clusters of items) can be in
practice interpreted as distinct ``dimensions'', similarly to
traditional latent factors - but without explicitly assuming their
existence (\citeproc{ref-christensen2021equivalency}{Christensen \&
Golino, 2021b}).

@@

For both studies, data analysis was carried out using R (4.4.3,
\citeproc{ref-Rcoreteam}{R Core Team, 2025}), leveraging the
\emph{easystats} ecosystem (\citeproc{ref-ludecke2019insight}{Lüdecke et
al., 2019}, \citeproc{ref-ludecke2020extracting}{2020}) and the
\emph{EGAnet} package
(\citeproc{ref-christensen2021estimating}{Christensen \& Golino,
2021a}). The reproducible code and additional details are available at
https://github.com/RealityBending/InteroceptionScale.

@@

Next, we tested the predictive power of the Mint scale relative to other
interoception questionnaires. We assessed the importance of each
interoceptive dimension (from all the scales) as a unique predictor by
fitting regression models (linear for continuous measures - e.g.,
depression score - and logistic for binary variables - e.g., presence
\emph{vs.} absence of mood disorders) to predict different outcomes, and
compare the best between the 4 interoception scales based on the total
explained variance (R2 or pseudo-R2 for logistic models,
\citeproc{ref-ludecke2021performance}{Lüdecke et al., 2021}). We then
evaluated the predictive performance of each scale as a whole by
comparing regression models with all the dimensions entered as
predictors (note that the IAS and BPQ only have one total score
variable). We assessed the models based on their R2, as well as on the
Bayesian Information Criterion (BIC), which penalizes for predictor
number, thus offering a balance between model parsimony and predictive
power. For the logistic models, we quantified the discriminative power
by computing the Area Under the Curve (AUC) of the Receiver Operating
Characteristic (ROC) curve, which assesses the model's discriminative
power (the combination of sensitivity and specificity).

@@

\bibitem[\citeproctext]{ref-christensen2021estimating}
Christensen, A. P., \& Golino, H. (2021a). Estimating the stability of
psychological dimensions via bootstrap exploratory graph analysis: A
monte carlo simulation and tutorial. \emph{Psych}, \emph{3}(3),
479--500.

\bibitem[\citeproctext]{ref-christensen2021equivalency}
Christensen, A. P., \& Golino, H. (2021b). On the equivalency of factor
and network loadings. \emph{Behavior Research Methods}, \emph{53}(4),
1563--1580.

@@

\bibitem[\citeproctext]{ref-ludecke2020extracting}
Lüdecke, D., Ben-Shachar, M. S., Patil, I., \& Makowski, D. (2020).
Extracting, computing and exploring the parameters of statistical models
using r. \emph{Journal of Open Source Software}, \emph{5}(53), 2445.

\bibitem[\citeproctext]{ref-ludecke2021performance}
Lüdecke, D., Ben-Shachar, M. S., Patil, I., Waggoner, P., \& Makowski,
D. (2021). Performance: An r package for assessment, comparison and
testing of statistical models. \emph{Journal of Open Source Software},
\emph{6}(60).

\bibitem[\citeproctext]{ref-ludecke2019insight}
Lüdecke, D., Waggoner, P. D., \& Makowski, D. (2019). Insight: A unified
interface to access information from model objects in r. \emph{Journal
of Open Source Software}, \emph{4}(38), 1412.

@@

\bibitem[\citeproctext]{ref-Rcoreteam}
R Core Team. (2025). \emph{R: A language and environment for statistical
computing}. R Foundation for Statistical Computing.
\url{https://www.R-project.org/}
